Re-Allocation of Resources during Releases \titlenote{Copyright note}

### 2013-10-30 paper1/paper1.tex
Ownership calculated for releases. Paper updated. Reference files renamed.

\title{Re-Allocation of Resources during Releases \titlenote{Copyright note}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development; Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{1_brooks}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR effects project organization and introduces technical debt. The key research questions that we expect to answer with our methodology are as follows:
\begin{enumerate}
\item Do developers work on different areas of the system around the time of release?
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)?
\item Do the areas of code that are modified around the time of release have higher defect densities than code that is modified during normal development?
\end{enumerate}

We observe a reallocation of the resources among the software development teams in a large project to identify that an improper reallocation or inappropriate reorganization causes a disruptive event take place in a software development process. We attempt to identify a project’s different release times and calculate the difference between two consecutive releases to discover which are the new areas have been worked between two consecutive releases. The commit log data that we are working on for this purpose will help us to extract a lot of information like calculating the developers' working areas and time-frame of each release. This information will help us to identify the criteria of the resources, their roles, file ownership and nativeness (∆ƞ) in the domain. This knowledge of nativeness (∆ƞ) will guide us understanding the method of reallocation.
Very few research works have been performed regarding the re-allocation of resources.  Robert van Engelen worked for similar kind of a research to understand the resource allocation dynamics across the software projects \cite{2_datta}. He mainly tried to reallocate development resources amongst projects for increasing the satisfactory level of consumer or customer while we are focusing on the impact on code-base like the complexity of script files. Robert proposed a project-entropy metric in his work to understand if there is any limit for a particular reallocation does not lead to user satisfaction. Here entropy is to represent disorder and chaos to understand degradation of software and its inherent complexity. In his work resources may not just be the developers but also can be any other resources necessary for a software project development.
We have organized this paper as follows. In Section 2, we describe some background and motivations followed by some summaries of related works in section 3. Section 4 will describe about the ownership of files and ownership of a set of files or a directories. We will try to understand how native a code-base is to a developer or a development team. In Section 5, some analysis to determine reallocation has been performed in a release or where reallocation needs to be performed will be presented. What changes in nativeness (∆ƞ) occurs after the reallocation. Section 6 will give us the result to show how change in ∆ƞ puts impact on the outcome of a software. Finally section 7 will give us an idea of our future work followed by the conclusion in section 8.

@@

\section{Methodology and Data}
This section presents our methodology for discovering information which can give us the idea to get the answers to our research questions. We have collected the development history data of Linux kernel. Actually it is a database containing the all the commit log records by the Linux kernel developers since 2005. We are going to present the steps involved in this process and then we will follow up with an application of our methodology in a case study.
Our methodology can be summarized as: Extracting Data for revisions and releases (Section 4.1); Partitioning the version numbers (Section 4.2); Get time-span between each release (Section 4.3); Calculate developer areas (Section 4.4); Finding code area owners (Section 4.5). Finding merging time and development time within a release period that we found in section 4.1 (Section 4.6).

@@

\subsection{Partitioning Release Numbers}
We stored the git commit log extracted data into a table named git commit and git revision where all the basic and log information for a particular commit was mentioned in the first table and second one containing which commit belongs to which version of Linux kernel development and change details like path modified, new path created due to the change, how many addition and how many deletion occurred in a particular commit etc. By joining these tables we can easily get the dates of each version and from the version number which is a combination of different types of releases we can get determine which commit belongs to which version and release, and also what type of release that is as well. Figure 1 shows the picture of the data how it looks like.
\begin{figure}
\centering
\psfig{file=versions.png, height=1in, width=3in,}
\caption{Releases extracted from Version}
\end{figure}

@@

Note that the very first row has no information for the ``Start Date'' because we don't have any data prior to release ``linuxv2.6.12''. From the extraction we found complete records for releases ``linuxv2.6.13'' to ``linuxv2.6.39'' then ``linuxv3.0'' to ``linuxv3.11''. We have last couple of release candidates and the release date for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12'' which is not useful for our calculation. So we are having 340 releases including major, minor, micro and rc; 40 complete releases from ``linuxv2.6.12'' to ``linuxv3.11''. We have total 400441 distinct commits and for 372943 of them have change information for 77082 different files in our hand where we have 14599 developers' working history among 14621 distinct developers.

@@

\subsection{Finding Code-Area Owners}
... ... ...

@@

>>>>

\title{Re-Allocation of Resources during Releases \titlenote{Copyright note}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development, Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Code Ownership, Defect Density, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{1_brooks}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR effects project organization and introduces technical debt. The key research questions that we expect to answer with our methodology are as follows:
\begin{enumerate}
\item Do developers work on different areas of the system around the time of release?
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)?
\item Do the areas of code that are modified around the time of release have higher defect densities than code that is modified during normal development?
\end{enumerate}

We observe a reallocation of the resources among the software development teams in a large project to identify that an improper reallocation or inappropriate reorganization causes a disruptive event take place in a software development process. We attempt to identify a project’s different release times and calculate the difference between two consecutive releases to discover which are the new areas have been worked between two consecutive releases. The commit log data that we are working on for this purpose will help us to extract a lot of information like calculating the developers' working areas and time-frame of each release. This information will help us to identify the criteria of the resources, their roles, file ownership and nativeness in the domain. This knowledge of nativeness will guide us understanding the method of reallocation.
Very few research works have been performed regarding the re-allocation of resources.  Robert van Engelen worked for similar kind of a research to understand the resource allocation dynamics across the software projects \cite{2_datta}. He mainly tried to reallocate development resources amongst projects for increasing the satisfactory level of consumer or customer while we are focusing on the impact on code-base like the complexity of script files. Robert proposed a project-entropy metric in his work to understand if there is any limit for a particular reallocation does not lead to user satisfaction. Here entropy is to represent disorder and chaos to understand degradation of software and its inherent complexity. In his work resources may not just be the developers but also can be any other resources necessary for a software project development.
We have organized this paper as follows. In Section 2, we describe some background and motivations followed by some summaries of related works in section 3. Section 4 will describe about the ownership of files and ownership of a set of files or a directories. We will try to understand how native a code-base is to a developer or a development team. In Section 5, some analysis to determine reallocation has been performed in a release or where reallocation needs to be performed will be presented. What changes in nativeness (∆ƞ) occurs after the reallocation. Section 6 will give us the result to show how change in nativeness puts impact on the outcome of a software. Finally section 7 will give us an idea of our future work followed by the conclusion in section 8.

@@

\section{Methodology and Data}
This section presents our methodology for discovering information which can give us the idea to get the answers to our research questions. We have collected the development history data of Linux kernel. Actually it is a database containing the all the commit log records by the Linux kernel developers since 2005. We are going to present the steps involved in this process and then we will follow up with an application of our methodology in a case study. In order to address our research questions, we obtain key measures of project evolution from archival data had been preserved throughout the history of the Linux Kernel development project. All information on the OSS project is recorded in electronic form. Many other OSS projects archive similar data, so the techniques used here can be replicated on any such project. We used the data elements extracted from the archival source to construct a number of measures on the commit log records to understand the behavior of the development resources (mainly developers).
Our methodology can be summarized as: Extracting Data for revisions and releases (Section 4.1); Partitioning the version numbers (Section 4.2); Get time-span between each release (Section 4.3); Calculate developer areas (Section 4.4); Finding code ownership (Section 4.5). Calculating Jaccard Distance for different development areas.

@@

\subsection{Partitioning Release Numbers}
We stored the git commit log extracted data into a table named git commit and git revision where all the basic and log information for a particular commit was mentioned in the first table and second one containing which commit belongs to which version of Linux kernel development and change details like path modified, new path created due to the change, how many addition and how many deletion occurred in a particular commit etc. By joining these tables we can easily get the dates of each version and from the version number which is a combination of different types of releases we can get determine which commit belongs to which version and release, and also what type of release that is as well.

@@

Note that the very first row has no information for the ``Start Date'' because we don't have any data prior to stable release ``linuxv2.6.12''. From the extraction we found complete records for releases ``linuxv2.6.13'' to ``linuxv2.6.39'' then ``linuxv3.0'' to ``linuxv3.11''. We have last couple of release candidates and the release date for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12'' which is not useful for our calculation. So we are having 340 releases including major, minor, micro and rc; 40 stable releases from ``linuxv2.6.12'' to ``linuxv3.11''. We have total 400441 distinct commits and for 372943 of them have change information for 77082 different files in our hand where we have 14599 developers' working history among 14621 distinct developers.
Surprisingly 14569 individual files didn't have any change (addition or deletion) but was being committed at least once. So we found them in our released wise datasets.

@@

These information above will help us to answer our first research-question.

\subsection{Finding Code Ownership}
So many developers are working from different development communities. We tried to investigate the "code ownership" \cite{15_mockus} evolved in Linux Kernel development. Here code ownership is going to be represented as percentage value. Considering the data that we have for complete releases (``linuxv2.6.12'' to ``linuxv3.11'') we filtered the code files and we found 75138 distinct files that have been worked on. We investigate the total number of commits, total number of churn for each and every file also how many developers contributed for these commits and changes. This information yet does not tell about the ownership. We need to find out ownership for every developers in different development areas in an RP. We already have the collection of developers working in different areas MP, RDP, RTR within an RP and stored them with corresponding file they have worked on, how many commits a developer made to a particular file and how many changes he/she has made. We now update those information with the ownership information. For calculating ownership Equation 1 is applied.
\begin{equation}\omega=\frac{n}{N}*100\end{equation}
Where $\omega$ is the ownership and n is the number of total changes made by an atuthor of a file, N is the total number of changes made for a particular file.
To determine a developer be a owner of a file we considering that if the developer makes changes more than 80\% of the total change that the file has got, then that developer can be treated as a owner of the codes of that file similarly as C. Gutwin did \cite{16_gutwin} for finding out the main developers' group for a project. We already mentioned that 14569 individual files didn't have any change (addition or deletion) but was being committed at least once. So we found them in our release wise datasets that's why in addition to Gutwin here considering commits also to determine ownership but only for files without any add or remove lines information. If a file has no addition or deletion information (we are calling it churn) in it's commit log but has beet committed by some one at least once then that or those authors may have the ownership for that file. So the equation 1 becomes like:
\begin{equation} \omega = \left\{ \begin{array}{l l} n/N & \quad \text{if $N > 0$}\\ c/C & \quad \text{if $N = 0$} \end{array} * 100 \right.\end{equation}
Where c is the number of commits made by the author and C is the total number of commits made for the particular file.
We calculate code ownership for all different parts MP, RDP, RTR and also for an entire release period RP. Table 6 represents a part of the data where ownerships is calculated.

\begin{table}[ht]
\caption{Code Ownership in RP}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Author 				& Linuxv		& Path				& $\omega$ (\%) \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
Thomas Gleixner		& 2.6.24		& .../numa\_64.h	& 0.0000\\
Jeff Kirsher			& 3.1		& .../ethtool.c		& 0.0000\\
Sathya Perla			& 2.6.29		& .../hwlib.h		& 100.00\\
Roel Kluin			& 2.6.24		& .../innovator.h 	& 100.00\\
David S. Miller		& 2.6.24		& .../visasm.h 		& 100.00\\
Stephen Hemminger	& 2.6.24		& .../qla3xxx.c	 	& 12.083\\
Randy Dunlap			& 2.6.24		& ...\_core.c 		& 100.00\\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}
The example given above is calculated for the developers worked for different files in different releases. Ownership is calculated for all the developers in release periods. In the same way we have calculated what is the ownership value for a developer to a particular file in the merge period (MP) in a release, development period (RDP) in a release and RTR period ind a release.

@@

\section{Preliminary Results}
In this section we present results from several quantitative analysis of the archival data from the Linux Kernel development project. The measures we derive from these data are well-suited to address our research questions.
